---
title: "Assignment 1"
author: "Emma Siegfried"
date: "May 2, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Library:
```{r message=FALSE}

library(tidyverse)
library(lubridate)
library(knitr)
library(kableExtra)

```


###Import and Tidy

1. Read the gazetteer data as-is (all columns; no type conversion) into a gaz_raw tibble. 
  
2. Copy only specified columns into a gaz tibble.
  
```{r results='hide', message=FALSE}

#Raw Dataset:
gaz_raw <- read_delim("CA_Features_20180401.txt", delim = "|")

#class(gaz_raw) #confirmed this is a tibble

#Subset data
gaz <- gaz_raw %>% 
  select(FEATURE_ID, FEATURE_NAME, FEATURE_CLASS, STATE_ALPHA, COUNTY_NAME, 
         PRIM_LAT_DEC, PRIM_LONG_DEC, SOURCE_LAT_DEC, SOURCE_LONG_DEC, ELEV_IN_M, MAP_NAME, DATE_CREATED, DATE_EDITED)

```
  
3. Convert columns to appropriate type. Convert placeholders for unknown data to NA
  
```{r results='hide'}

gaz$DATE_CREATED <- as.Date(gaz$DATE_CREATED, "%m/%d/%Y") 
gaz$DATE_EDITED <- as.Date.character(gaz$DATE_EDITED, "%m/%d/%Y") 


```
   
4. Delete rows with: primary latitude and longitude unknown; and if not in California
   
*Tidy Dataset:*
```{r results='hide'}

gaz <- gaz %>% 
  filter(STATE_ALPHA == "CA") %>% 
  filter(PRIM_LAT_DEC != "NA") %>% 
  filter(PRIM_LONG_DEC != "NA")

```
   
5. Write tibble to CSV file with "|" delimiter
   
*Create CSV:*
```{r results='hide'}

write_delim(gaz, path = "gaz_Emma.csv", delim = "|")

```
   
---
   
---
   
   
###Analyze
   
     
*Question 1:* What is the most frequently occurring feature name?
```{r results='hide'}

gaz$FEATURE_NAME <- as.factor(gaz$FEATURE_NAME)
gaz$COUNTY_NAME <- as.factor(gaz$COUNTY_NAME)

gaz_name_max <- gaz %>% 
  group_by(FEATURE_NAME) %>% 
  count(FEATURE_NAME)

arrange(gaz_name_max, desc(n))

table_name <- kable(gaz_name_max, "html") %>% 
  kable_styling(bootstrap_options = "striped", "hover")

table_name

```

**Answer:** Church of Christ
  
---
  

*Question 2:* What is the least frequently occuring feature name?
```{r results='hide'}

unique(gaz$FEATURE_CLASS)
gaz$FEATURE_CLASS <- as.factor(gaz$FEATURE_CLASS)

gaz_class_min <- gaz %>% 
  count(FEATURE_CLASS) %>% 
  group_by(FEATURE_CLASS)

arrange(gaz_class_min, desc(-gaz_class_min$n))

```

**Answer:** Isthmus and Sea
  
---
   
*Question 3:* What is the approximate center point of each county?

```{r}

#Points
gaz_points <- data.frame(
  long = gaz$PRIM_LONG_DEC,
  lat = gaz$PRIM_LAT_DEC,
  names = gaz$FEATURE_NAME,
  county = gaz$COUNTY_NAME,
  feature = gaz$FEATURE_CLASS,
  stringsAsFactors = FALSE
)


gaz_centers <- gaz_points %>% 
  group_by(county) %>% 
  summarise(
    Longitude = mean(long, na.rm = TRUE),
    Latitude = mean(lat, na.rm = TRUE)) %>% 
  rename(County = county)

table_centers <- gaz_centers %>% 
  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

table_centers

  
```
   
     
---
   
*Question 4:* What are the fractions of the total number of features in each county that are natural? man-made?
   
```{r message=FALSE, warning=FALSE}

#Raw Dataset:
features_raw <- read_csv("Class_Code_Definitions.csv") %>% 
  select(1) %>% 
  setNames('FEATURE_CLASS')

class_combo <- as.tibble(gaz_class_min)

class_combo <- class_combo %>% 
  full_join(features_raw, by = "FEATURE_CLASS") %>% 
  arrange(-desc(FEATURE_CLASS))

definition <- c("M","N","N","N","N","N","N","N","N","N","M","M","M","N","N","M","M","N","M","M","N","N","M","M","N","N", "N","N","N","N","M","M","N","N","N","N","MN","M","M","M","M","M","N","N","M","M","N","N","M","M","N","M","N","N","N","N", "N","N","M","M","M","NA","N","M","N")

class_combo <- add_column(class_combo, Type = definition)
  
gaz_join <- gaz %>% 
  full_join(class_combo, by = "FEATURE_CLASS")

gaz_join$COUNTY_NAME <- as.factor(gaz_join$COUNTY_NAME)

gaz_join_count <- gaz_join %>% 
  group_by(COUNTY_NAME) %>% 
  summarise(
    Manmade_Per_County = sum(Type == "M", na.rm = TRUE),
    Natural_Per_County = sum(Type == "N", na.rm = TRUE))

gaz_points_count <- gaz %>% 
  group_by(COUNTY_NAME) %>% 
  count(FEATURE_CLASS)

gaz_points_count <- gaz_points_count %>% 
  group_by(COUNTY_NAME) %>% 
  summarise(
    sum(n)
  )

gaz_points_count <- rename(gaz_points_count, Total_Per_County = "sum(n)")


gaz_join_count <- gaz_join_count %>% 
  full_join(gaz_points_count, by = "COUNTY_NAME")

gaz_join_fraction <- gaz_join_count %>% 
  group_by(COUNTY_NAME) %>% 
  summarise(
    Fraction_Manmade = Manmade_Per_County/Total_Per_County, na.rm = TRUE,
    Fraction_Natural = Natural_Per_County/Total_Per_County, na.rm = TRUE)

gaz_join_fraction <- gaz_join_fraction %>% 
  full_join(gaz_join_count, by = "COUNTY_NAME")


gaz_join_fraction <- gaz_join_fraction %>% 
  select(-3) %>% 
  rename(County = COUNTY_NAME)

fraction_table <- gaz_join_fraction %>% 
  kable("html") %>% 
  kable_styling(bootstrap_options = c("striped", "hover"))

fraction_table



```
   








