---
title: "ESM 262: Assignment 2"
author: "Emma Siegfried"
date: "May 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

###SetUp
Library:
```{r message=FALSE}

library(tidyverse)
library(lubridate)
library(knitr)
library(kableExtra)
library(dplyr)
library(DBI)

```


###Import and Tidy

1. Read the gazetteer data as-is (all columns; no type conversion) into a gaz_raw tibble. 
  
2. Copy only specified columns into a gaz tibble.
  
```{r results='hide'}

#Raw Dataset:
gaz_raw <-
  read_delim(
    "CA_Features_20180401.txt",
    delim = "|",
    col_types = cols(.default = col_character())
  )
#class(gaz_raw) #confirmed this is a tibble
```

```{r results='hide'}

#Subset data
gaz <- 
  select(gaz_raw,
    id         = FEATURE_ID,
    name       = FEATURE_NAME,
    class      = FEATURE_CLASS,
    state      = STATE_ALPHA,
    county     = COUNTY_NAME,
    lat        = PRIM_LAT_DEC,
    lon        = PRIM_LONG_DEC,
    src_lat    = SOURCE_LAT_DEC,
    src_lon    = SOURCE_LONG_DEC,
    elev       = ELEV_IN_M,
    map        = MAP_NAME,
    created_on = DATE_CREATED,
    edited_on  = DATE_EDITED
  )

```
  
3. Convert columns to appropriate type. Convert placeholders for unknown data to NA
 
*Tidy Dataset:*
```{r results='hide'}

gaz$id         = parse_integer(gaz$id)
gaz$lat        = parse_double(gaz$lat, na = "0")
gaz$lon        = parse_double(gaz$lon, na = "0")
gaz$src_lat    = parse_double(gaz$src_lat)
gaz$src_lon    = parse_double(gaz$src_lon)
gaz$elev       = parse_integer(gaz$elev)
gaz$map        = parse_character(gaz$map, na = "Unknown")
gaz$created_on = parse_date(gaz$created_on, format = "%m/%d/%Y")
gaz$edited_on  = parse_date(gaz$edited_on, format = "%m/%d/%Y")

```
   
4. Delete rows with: primary latitude and longitude unknown; and if not in California
```{r}

gaz <-
  filter(gaz,
    !is.na(lon) &
    !is.na(lat) &
    !is.na(county) &
    state == "CA"
  )

gaz

```

5. Create SQL database connection
```{r}


con <- DBI::dbConnect(RSQLite::SQLite(), path = ":memory:")

copy_to(con, gaz, "gaz",
        temporary = FALSE, 
        overwrite = TRUE, 
        indexes = list(
         "name"
        ))

#Create a table cookie that connects to the database

gaz_db <- tbl(con, "gaz")

```

---
  
###Analyze and Answer

**Q1. What is the most frequently occurring feature name?**
```{r}

count_names_max <- dbGetQuery(con, 
                          "SELECT NAME, COUNT() AS n 
                          FROM gaz GROUP BY NAME
                          ORDER BY n DESC
                          LIMIT 1")

count_names_max

```
**Answer:** Church of Christ appears the most with 228 listings

---

**Q2. What is the least-frequently occurring feature class?**
```{r}


count_class_min <- dbGetQuery(con,
                            "SELECT CLASS, COUNT () AS n
                            FROM gaz
                            GROUP BY CLASS
                            ORDER BY n
                            LIMIT 2")

count_class_min

```
**Answer:** Sea and Isthum appear the least with 1 listing each.

---

**Q3. What is the approximate center point of each count?**

```{r}


#Summarize Coordinates by County
gaz_centers <- dbGetQuery(con, 
        "SELECT COUNTY, 
                Center_Longitude, Center_Latitude 
         FROM (SELECT COUNTY, 
                MIN(LON) + MAX(LON) AS Center_Longitude,
                MIN(LAT) + MAX(LAT) AS Center_Latitude 
                FROM gaz
                GROUP BY COUNTY)")

head(gaz_centers)

```

**Answer**: County centers are listed in the gaz_centers query.

---

**Q4. What are the fractions of the total number of features in each county that are natural vs. manmade?**

Step 1: Create New Category Table

```{r catebory_tibble}

categories <- tribble(
  ~class,             ~category,
  "Airport",          "manmade",
  "Arch",             "natural",
  "Area",             "natural",
  "Arroyo",           "natural",
  "Bar",              "natural",
  "Basin",            "natural",
  "Bay",              "natural",
  "Beach",            "natural",
  "Bench",            "natural",
  "Bend",             "natural",
  "Bridge",           "manmade",
  "Building",         "manmade",
  "Canal",            "manmade",
  "Cape",             "natural",
  "Cemetery",         "manmade",
  "Census",           "manmade",
  "Channel",          "natural",
  "Church",           "manmade",
  "Civil",            "manmade",
  "Cliff",            "natural",
  "Crater",           "natural",
  "Crossing",         "manmade",
  "Dam",              "manmade",
  "Falls",            "natural",
  "Flat",             "natural",
  "Forest",           "natural",
  "Gap",              "natural",
  "Glacier",          "natural",
  "Gut",              "natural",
  "Harbor",           "manmade",
  "Hospital",         "manmade",
  "Island",           "natural",
  "Isthmus",          "natural",
  "Lake",             "natural",
  "Lava",             "natural",
  "Levee",            NA,
  "Locale",           "manmade",
  "Military",         "manmade",
  "Mine",             "manmade",
  "Oilfield",         "manmade",
  "Park",             "manmade",
  "Pillar",           "natural",
  "Plain",            "natural",
  "Populated Place",  "manmade",
  "Post Office",      "manmade",
  "Range",            "natural",
  "Rapids",           "natural",
  "Reserve",          "manmade",
  "Reservoir",        "manmade",
  "Ridge",            "natural",
  "School",           "manmade",
  "Sea",              "natural",
  "Slope",            "natural",
  "Spring",           "natural",
  "Stream",           "natural",
  "Summit",           "natural",
  "Swamp",            "natural",
  "Tower",            "manmade",
  "Trail",            "manmade",
  "Tunnel",           "manmade",
  "Valley",           "natural",
  "Well",             "manmade",
  "Woods",            "natural"
)
```


Step 2: Copy Categories Table to SQL Database
```{r}

#Copy into Database & Wrote out Table

copy_to(con, categories, "categories",
        temporary = FALSE, 
        overwrite = TRUE)

```


Step 3: Make smaller Gaz table
```{r}

#Gaz Table
gaz_mini <- dbGetQuery(con,
      "SELECT COUNTY, CLASS
       FROM gaz")

```


Step 4: Join Categories with Gaz
```{r}

#Join Categories and Gaz Tables

category_combo <- dbGetQuery(con, 
     "SELECT gaz.COUNTY, gaz.CLASS,  
             categories.CATEGORY
      FROM gaz 
      JOIN categories
      ON gaz.CLASS = categories.CLASS
      WHERE COUNTY IS NOT NULL");

```


Step 5: Write out Combined Table
```{r}

#Save combo table to database
copy_to(con, category_combo, "category_combo",
        temporary = FALSE, 
        overwrite = TRUE)

gaz_db_category <- tbl(con, "category_combo")

```


Step 6: Calculate Manmade & Natural fractions per County
```{r}

#Create Table of Feature Counts per County
category_fractions <- dbGetQuery(con,
     "SELECT COUNTY, 
      SUM(category = 'manmade') AS Manmade_Per_County,
      SUM(category = 'natural') AS Natural_Per_County,
      COUNT(category) AS Total_Per_County,
      ROUND(CAST(SUM(category = 'manmade') AS REAL)/COUNT(category),2) 
          AS Fraction_Manmade,
      ROUND(CAST(SUM(category = 'natural') AS REAL)/COUNT(category),2)
          AS Fraction_Natural
      FROM category_combo
      GROUP BY County")

head(category_fractions)

```


Step 7: Write out Fraction Table
```{r}

#Save fraction table to database
copy_to(con, category_fractions, "category_fractions",
        temporary = FALSE, 
        overwrite = TRUE)

gaz_db_fractions <- tbl(con, "category_fractions")

```



